---
title: "Homework4"
author: "Trevor Caughlin"
date: "February 6, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1. Gamma distribution
Explore the properties of the gamma distribution using the probability density function in R:

```{r gamma}
args(dgamma)
```

Use the `curve` function to explore how the shape and rate parameters alter the likelihood function. As output, include (1) a brief written explanation of how the shape and rate parameters determine the behavior of the function and (2) several different curves showing different parameter values for `dgamma` (For an example of similar code, see the beta distribution section of lecture 9).

Question 2. Find a published paper in your research area that uses either gamma or beta regression. Briefly (in one paragraph) answer the following questions:

What were the response and predictor variables? 
How were the results presented?
Do you agree with the interpretation of the results?

If you have trouble finding a paper that uses one of these analyses, here are several choices:

http://onlinelibrary.wiley.com/doi/10.1111/mec.12059/full
http://onlinelibrary.wiley.com/doi/10.1111/1365-2664.12531/full
http://onlinelibrary.wiley.com/doi/10.1002/rse2.33/full

Question 3: Apply beta or gamma regression to your own data (or a dataset you find on data dryad). Note that the notation for gamma and beta regression is slightly different than the other models we have covered:

```{r gamglm}
#beta regression:
library("betareg")
#beta_model=betareg(response~predictor)

#gamma regression:
#gamma_model=glm(response~predictor,family=Gamma(link="log"))

```

As output, include (1) a plot with model predictions overlaid on raw data, (2) a verbal description of effect size, and (3) estimates of significance, using `confint`

Question 4: Simulate a gamma-distributed random variable that is related to a predictor variable using `rgamma`. Describe how you reparameterized R's default gamma parameters to represent a mean value. Include a plot of your response vs. predictor variable, and a description of a biological process your data could represent.

Hint:
```{r gamsim}
#y <- rgamma(n, rate = shape / y_mean, shape = shape)
```
